---
title: "Project 1"
author: "Aspen Saplan"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
library(ezids)
```

# Set up of dataset walmart:
```{r, echo=TRUE}
#getwd()
walmart = read.csv("Walmart.csv")
walmart$category = as.factor(walmart$category)
walmart$store_location = as.factor(walmart$store_location)
walmart$customer_loyalty_level = as.factor(walmart$customer_loyalty_level)
walmart$payment_method = as.factor(walmart$payment_method)
print(str(walmart))
```

# Summaries of variables:
### The summary of category:
```{r,echo=TRUE}
summary(walmart$category)
```

### The summary of store_location:
```{r,echo=TRUE}
summary(walmart$store_location)
```

### The summary of customer_income:
```{r,echo=TRUE}
summary(walmart$customer_income)
```

### The summary of customer_loyalty_level:
```{r,echo=TRUE}
summary(walmart$customer_loyalty_level)
```

### The summary of payment_method:
```{r,echo=TRUE}
summary(walmart$payment_method)
```

# Checking if NA values exist:

### Category
Sum of category NA values: `r sum(is.na(walmart$category))`

### store_location
Sum of store_location NA values: `r sum(is.na(walmart$store_location))`

### customer_income
Sum of customer_income NA values: `r sum(is.na(walmart$customer_income))`

### customer_loyalty_level
Sum of customer_loyalty_level NA values: `r sum(is.na(walmart$customer_loyalty_level))`

### payment_method
Sum of payment_method NA values: `r sum(is.na(walmart$payment_method))`

# EDA of category and store_location:
```{r,echo=TRUE}
library(ggplot2)
ggplot(walmart, aes(x = store_location, fill = category)) +
  geom_bar(position = "dodge", color = "black")+
  scale_fill_manual(values = c(
    "Appliances" = "purple",
    "Electronics" = "red"
  ))+
  labs(title = "Number of Transactions by Store Location and Category",
    x = "Store Location",
    y = "Number of Transactions",
    fill = "Category")
```

# EDA of category and customer_income:
```{r,echo=TRUE}
ggplot(walmart, aes(x = category, y=customer_income)) + 
  geom_boxplot() + 
  geom_boxplot( color="black", fill="gold", outlier.colour="blue", outlier.shape=19, outlier.size=2) +
  labs(title="Box plot of Customer Income in Walmart using `ggplot`", x="", y = "Customer Income(Dollar Amount)")
```

The data looks pretty similar across both appliances and electronics.

```{r, echo=TRUE}
library(ggplot2)
ggplot(data=walmart, aes(customer_income)) + 
  geom_histogram(breaks=seq(20005.34, 119999.8, by = 2000),
                 col="black", 
                 fill="gold", 
                 alpha = .8) + # opacity
  labs(x="Customer Income(Dollar Amount)", y="Frequency") +
  labs(title="Histogram of Customer Income for Walmart Dataset using `ggplot`")
```

```{r, echo=TRUE}
qqnorm(walmart$customer_income, main="Q-Q plot of Customer Income in Walmart") 
qqline(walmart$customer_income)
```

The distribution for the customer_income as a whole is approximately uniform with no strong skewing.

## Subset of appliances:
```{r, echo=TRUE}
appliances = subset(walmart, category == "Appliances")
print(head(appliances[, c("category", "customer_income")], n=5))
```
```{r, echo=TRUE}
library(ggplot2)
ggplot(data=appliances, aes(customer_income)) + 
  geom_histogram(breaks=seq(20005.34, 119999.8, by = 2000),
                 col="black", 
                 fill="red", 
                 alpha = .8) + # opacity
  labs(x="Customer Income(Dollar Amount)", y="Frequency") +
  labs(title="Histogram of Customer Income for Appliance Dataset using `ggplot`")
```

```{r, echo=TRUE}
qqnorm(appliances$customer_income, main="Q-Q plot of Customer Income in Appliances") 
qqline(appliances$customer_income)
```

The distribution for the customer_income for appliances is approximately uniform with no strong skewing.

## Subset of electronics:
```{r, echo=TRUE}
electronics = subset(walmart, category == "Electronics")
print(head(electronics[, c("category", "customer_income")], n=5))
```

```{r, echo=TRUE}
library(ggplot2)
ggplot(data=electronics, aes(customer_income)) + 
  geom_histogram(breaks=seq(20028.81, 119997.5, by = 2000),
                 col="black", 
                 fill="green", 
                 alpha = .8) + # opacity
  labs(x="Customer Income(Dollar Amount)", y="Frequency") +
  labs(title="Histogram of Customer Income for Electronics Dataset using `ggplot`")
```

```{r, echo=TRUE}
qqnorm(electronics$customer_income, main="Q-Q plot of Customer Income in Electronics") 
qqline(electronics$customer_income)
```

The distribution for the customer_income for electronics is approximately uniform with no strong skewing.

# EDA of category and customer_loyalty_level:
```{r,echo=TRUE}
library(ggplot2)
ggplot(walmart, aes(x = customer_loyalty_level, fill = category)) +
  geom_bar(position = "dodge", color = "black")+
  scale_fill_manual(values = c(
    "Appliances" = "purple",
    "Electronics" = "red"
  ))+
  labs(title = "Number of Transactions by Customer Loyalty Level and Category",
    x = "Customer Loyalty Level",
    y = "Number of Transactions",
    fill = "Category")
```

# EDA of category and payment_method:
```{r,echo=TRUE}
library(ggplot2)
ggplot(walmart, aes(x = payment_method, fill = category)) +
  geom_bar(position = "dodge", color = "black")+
  scale_fill_manual(values = c(
    "Appliances" = "purple",
    "Electronics" = "red"
  ))+
  labs(title = "Number of Transactions by Payment Method and Category",
    x = "Payment Method",
    y = "Number of Transactions",
    fill = "Category")
```

# Chi Square Test of store_location and category
```{r,echo=TRUE}
contableSL = table(walmart$category, walmart$store_location)
xkabledply(contableSL, title="Contingency table for category and store_location within walmart dataset")
```

## Test of Independence on category and store_location on contable

Null Hypothesis: Category and store_location are independent

Alternative Hypothesis: Category and store_location are not independent

Alpha = 0.05

```{r, echo=TRUE}
chitestSL = chisq.test(contableSL)
chitestSL
```
Conclusion: Since the p-value(0.1489722) is greater than the alpha(0.05) we can conclude that we cannot reject the null hypothesis.  Which provides evidence suggesting that store_location and category are independent.

# T-Test of customer_income and category
Customer_income Null Hypothesis: means are equal for appliances and electronics

Customer_income Alternative Hypothesis: means are not equal for appliances and electronics

Alpha = 0.05

```{r,echo=TRUE}
ttestCI = t.test(appliances$customer_income, electronics$customer_income)
ttestCI
```

Conclusion: Since the p-value(0.1725206) is less than the alpha(0.05) we can conclude by failing to reject the null hypothesis.  Which provides evidence suggesting that the mean value of customer_income for appliances is equal to electronics

# Chi Square Test of customer_loyalty_level and category
```{r,echo=TRUE}
contableCLL = table(walmart$category, walmart$customer_loyalty_level)
xkabledply(contableCLL, title="Contingency table for category and customer_loyalty_level within walmart dataset")
```

## Test of Independence on category and customer_loyalty_level on contable

Null Hypothesis: Category and customer_loyalty_level are independent

Alternative Hypothesis: Category and customer_loyalty_level are not independent

Alpha = 0.05

```{r, echo=TRUE}
chitestCLL = chisq.test(contableCLL)
chitestCLL
```

Conclusion: Since the p-value(0.3872087) is greater than the alpha(0.05) we can conclude that we cannot reject the null hypothesis.  Which provides evidence suggesting that customer_loyalty_level and category are independent.

# Chi Square Test of payment_method and category
```{r,echo=TRUE}
contablePM = table(walmart$category, walmart$payment_method)
xkabledply(contablePM, title="Contingency table for category and payment_method within walmart dataset")
```
## Test of Independence on category and payment_method on contable

Null Hypothesis: Category and payment_method are independent

Alternative Hypothesis: Category and payment_method are not independent

Alpha = 0.05
```{r, echo=TRUE}
chitestPM = chisq.test(contablePM)
chitestPM
```
Conclusion: Since the p-value(0.8444539) is greater than the alpha(0.05) we can conclude that we cannot reject the null hypothesis.  Which provides evidence suggesting that payment_method and category are independent.
