---
title: "Walmart Sales Analysis: Electronics vs Appliances (EDA)"
author: "Pablo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
```{r, echo=TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
library(janitor)
library(effsize)
```



```{r, echo=TRUE}
walmart <- read_csv("Walmart.csv") |> clean_names()
glimpse(walmart)
```

```{r, echo=TRUE}
walmart$revenue <- walmart$quantity_sold * walmart$unit_price
```

```{r, echo=TRUE}
table(walmart$category)
prop.table(table(walmart$category))
```

```{r, echo=TRUE}
library(dplyr)
```

```{r, echo=TRUE}
walmart %>%
  group_by(category) %>%
  summarise(
    total_revenue = sum(revenue),
    mean_revenue = mean(revenue),
    median_revenue = median(revenue),
    sd_revenue = sd(revenue),
    mean_quantity = mean(quantity_sold),
    n_transactions = n()
  )
```

With this in mind now I create a boxplot of revenue by category to compare medians and visualize any outliers.

```{r, echo=TRUE}
ggplot(walmart, aes(x = category, y = revenue)) +
  geom_boxplot() +
  labs(title = "Revenue Distribution by Category")
```
```{r, echo=TRUE}
ggplot(walmart, aes(x = category, y = quantity_sold)) +
  geom_boxplot() +
  labs(title = "Quantity Sold Distribution by Category")
```

```{r, echo=TRUE}
walmart %>%
  group_by(category) %>%
  summarise(total_revenue = sum(revenue)) %>%
  ggplot(aes(x = category, y = total_revenue)) +
  geom_col() +
  labs(title = "Total Revenue by Category")
```
```{r, echo=TRUE}
t.test(revenue ~ category, data = walmart)
```

A Welch two-sample t-test was conducted to compare mean revenue per transaction between Electronics and Appliances. The test was not statistically significant (t = -0.60, p = 0.548). The 95% confidence interval for the difference in means (-171.88, 91.22) includes zero, suggesting no meaningful difference in average revenue between the two categories.


```{r, echo=TRUE}
t.test(quantity_sold ~ category, data = walmart)
```


```{r, echo=TRUE}
library(effsize)
cohen.d(revenue ~ category, data = walmart)
```