---
title: "Project 1"
author: "Aspen Saplan"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
# Walmart
```{r init, include=F}
library(ezids)
```

#### - S.M.A.R.T Questions:

Which customer, store, time, promotion, and weather factors are associated with whether a Walmart purchase is an appliance or an electronic product?

This Walmart dataset contains 5000 records of transactions.  Each individual record represents one transaction from a customer.  Each individual record has a variable attached to the transaction representing what type of item was purchased.  That variable is called the category variable and its values are either electronics or appliances.  There are many other variables within our dataset, but we limited our investigation down to 14 variables that could have an association with the category variable.  Our data set was gathered through the free online platform Kaggle.com.  Some analysis that is seen within real world retail companies are studies such as demographic of customers and their locations or seasonal sales trends and promotional effects.  Initially with the background of retail analysis we formatted our first try at the smart question based on aggregations such as total revenue and/or average revenue per transaction.
However, once we did our Exploratory Data Analysis we started to see these 14 variables could have associations with our category variable.  Which led to us abandoning our original smart question, and just working through the exploratory data analysis and the hypothesis testing of all 14 variables.  Once we started to see a more clear answer to our unknown question we were able to put what we were looking for on paper.  It was hard to initially say our smart question due to having so many variables to look at, but once we analyzed all the variables it was easy to formulate our final smart question.  Overall, our research is focussed on finding which individual variables have statistically shown evidence of an association with the variable category being either electronics or appliances.


# 1. Store Location, Customer Loyalty Level, Payment Method, and Customer Income On Category:

## Author of this section: "Aspen Saplan"

## Set up of dataset walmart:
```{r, echo=TRUE}
#getwd()
walmart = read.csv("Walmart.csv")
walmart$category = as.factor(walmart$category)
walmart$store_location = as.factor(walmart$store_location)
walmart$customer_loyalty_level = as.factor(walmart$customer_loyalty_level)
walmart$payment_method = as.factor(walmart$payment_method)
print(str(walmart$category))
print(str(walmart$store_location))
print(str(walmart$customer_loyalty_level))
print(str(walmart$payment_method))
print(str(walmart$customer_income))
```

The variables that will be focussed on in this section will be store location, customer loyalty level, payment method, and  customer income when comparing to the category variable. 

The three qualitative or factor variables in this section are store location, customer loyalty level, and payment method. Store location is made up of four total locations being Chicago, IL: Dallas, TX; Los Angeles, CA; Miami, FL; New York, NY. Customer loyalty level consists of four total levels being platinum, gold, silver and bronze.  Payment method consists of four total payment options being cast, credit card, debit card, and digital wallet.

#### - Summary statistics of store location:

```{r,echo=TRUE}
summary(walmart$store_location)
```

#### - Summary statistics of customer loyalty level:

```{r,echo=TRUE}
summary(walmart$customer_loyalty_level)
```

#### - Summary statistics of payment method:

```{r,echo=TRUE}
summary(walmart$payment_method)
```

## EDA:

For the following factor variables I used a grouped bar chart with the number of transactions as the Y axis and the variable as the X axis.  I then filled them by which category they were, either appliances or electronics.  

```{r,echo=TRUE}
library(ggplot2)
ggplot(walmart, aes(x = store_location, fill = category)) +
  geom_bar(position = "dodge", color = "black")+
  scale_fill_manual(values = c(
    "Appliances" = "purple",
    "Electronics" = "red"
  ))+
  labs(title = "Number of Transactions by Store Location and Filled with Category",
    x = "Store Location",
    y = "Number of Transactions",
    fill = "Category")
```

```{r,echo=TRUE}
library(ggplot2)
ggplot(walmart, aes(x = customer_loyalty_level, fill = category)) +
  geom_bar(position = "dodge", color = "black")+
  scale_fill_manual(values = c(
    "Appliances" = "purple",
    "Electronics" = "red"
  ))+
  labs(title = "Number of Transactions by Customer Loyalty Level and Filled with Category",
    x = "Customer Loyalty Level",
    y = "Number of Transactions",
    fill = "Category")
```

```{r,echo=TRUE}
library(ggplot2)
ggplot(walmart, aes(x = payment_method, fill = category)) +
  geom_bar(position = "dodge", color = "black")+
  scale_fill_manual(values = c(
    "Appliances" = "purple",
    "Electronics" = "red"
  ))+
  labs(title = "Number of Transactions by Payment Method and Filled with Category",
    x = "Payment Method",
    y = "Number of Transactions",
    fill = "Category")
```

Looking at the different grouped bar charts that consist of store location, customer loyalty level, and payment method comparing them to the number of transactions filled by the category variable.  The electronics transactions have a tendency to exceed the appliance transactions for the most part.  There are a few cases such as New York, NY having appliance transactions exceed electronics transactions or the silver customer loyalty level having the same amount of electronic transactions as appliance transactions.  So, there is a pattern of electronics out weighing appliances for when looking at store location, customer loyalty level, and payment method.

The last variable to be explored in this section is customer income.  This variable is numerical so no need to factor it as R will treat this variable correctly. However, I will split appliances and electronics transactions into subsets in order to see the distribution for each category by customer income.

#### - Summary of customer income:

```{r,echo=TRUE}
summary(walmart$customer_income)
```

```{r,echo=TRUE}
ggplot(walmart, aes(x = category, y=customer_income)) + 
  geom_boxplot() + 
  geom_boxplot( color="black", fill="gold", outlier.colour="blue", outlier.shape=19, outlier.size=2) +
  labs(title="Box plot of Customer Income in Walmart using `ggplot`", x="", y = "Customer Income(Dollar Amount)")
```

#### - Subset of appliances:

```{r, echo=TRUE}
appliances = subset(walmart, category == "Appliances")
print(head(appliances[, c("category", "customer_income")], n=5))
```

#### - Summary of appliances customer income:

```{r,echo=TRUE}
summary(appliances$customer_income)
```

```{r, echo=TRUE}
library(ggplot2)
ggplot(data=appliances, aes(customer_income)) + 
  geom_histogram(breaks=seq(20005.34, 119999.8, by = 2000),
                 col="black", 
                 fill="red", 
                 alpha = .8) + # opacity
  labs(x="Customer Income(Dollar Amount)", y="Frequency") +
  labs(title="Histogram of Customer Income for Appliance Dataset using `ggplot`")
```

#### - Subset of electronics:
```{r, echo=TRUE}
electronics = subset(walmart, category == "Electronics")
print(head(electronics[, c("category", "customer_income")], n=5))
```

#### - Summary of electronics customer income:

```{r,echo=TRUE}
summary(electronics$customer_income)
```

```{r, echo=TRUE}
library(ggplot2)
ggplot(data=electronics, aes(customer_income)) + 
  geom_histogram(breaks=seq(20028.81, 119997.5, by = 2000),
                 col="black", 
                 fill="green", 
                 alpha = .8) + # opacity
  labs(x="Customer Income(Dollar Amount)", y="Frequency") +
  labs(title="Histogram of Customer Income for Electronics Dataset using `ggplot`")
```

Looking at both electronics customer income and appliances customer income using a histogram, it suggests that both electronics and appliances have a pretty uniform distribution.  Looking at a boxplot of the two categories shows that they have pretty similar descriptive statistics.  Thus, through the descriptive statistics there is not a huge pattern or trend regarding customer income and category.  The distributions are very similar to each other with no skewing at any value.  While the descriptive statistics such as mean, median, and inter quartile ranges are also similar to each other indicating customer income may not have a huge impact on which category is purchased.

## Sketched answer:

My overall quick sketch of an answer to our question would be that customer income most likely will not have an association to the category variable, while the other categorical variables may have an association with the category variable.

## Hypothesis Testing:

A chi-squared test must be used for the store location, customer loyalty level, and payment method variable when testing for independence with the category variable.  A chi-squared test must be used due to store location, customer loyalty, and payment method being categorical variables and the category variable being categorical variable as well.  First a contingency table has to be made for store location and category, customer loyalty level and category, and payment method and category.

```{r,echo=TRUE}
contableSL = table(walmart$category, walmart$store_location)
xkabledply(contableSL, title="Contingency table for category and store_location within walmart dataset")
```

```{r,echo=TRUE}
contableCLL = table(walmart$category, walmart$customer_loyalty_level)
xkabledply(contableCLL, title="Contingency table for category and customer_loyalty_level within walmart dataset")
```

```{r,echo=TRUE}
contablePM = table(walmart$category, walmart$payment_method)
xkabledply(contablePM, title="Contingency table for category and payment_method within walmart dataset")
```

#### - Test of Independence on category and store_location using contableSL:

Null Hypothesis: Category and store_location are independent

Alternative Hypothesis: Category and store_location are not independent

Alpha = 0.05

```{r, echo=TRUE}
chitestSL = chisq.test(contableSL)
chitestSL
```

After running our chi-squared test on store location and category we can see that our p-value returned is 0.149.  Comparing our p-value(0.149) and our alpha value(0.05) we can see that the p-value is greater than alpha.  Thus, since our p-value is greater than our alpha value we fail to reject the null hypothesis, which provides evidence suggesting that store location and category are independent and have no association. 

#### - Test of Independence on category and customer_loyalty_level using contableCLL:

Null Hypothesis: Category and customer_loyalty_level are independent

Alternative Hypothesis: Category and customer_loyalty_level are not independent

Alpha = 0.05

```{r, echo=TRUE}
chitestCLL = chisq.test(contableCLL)
chitestCLL
```

After running our chi-squared test on customer loyalty level and category we can see that our p-value returned is 0.3872.  Comparing our p-value(0.3872) and our alpha value(0.05) we can see that the p-value is greater than alpha.  Thus, since our p-value is greater than our alpha value we fail to reject the null hypothesis, which provides evidence suggesting that customer loyalty level and category are independent and have no association. 

#### - Test of Independence on category and payment_method using contablePM:

Null Hypothesis: Category and payment_method are independent

Alternative Hypothesis: Category and payment_method are not independent

Alpha = 0.05
```{r, echo=TRUE}
chitestPM = chisq.test(contablePM)
chitestPM
```

After running our chi-squared test on payment method and category we can see that our p-value returned is 0.8445.  Comparing our p-value(0.8445) and our alpha value(0.05) we can see that the p-value is greater than alpha.  Thus, since our p-value is greater than our alpha value we fail to reject the null hypothesis, which provides evidence suggesting that payment method and category are independent and have no association.

For the customer income variable I started by separating appliance transactions and electronics transactions into two subsets.  Knowing that customer income is a numerical variable I know that a two sample t-test is needed for this hypothesis test.

#### - T-Test of customer_income and category:

Customer_income Null Hypothesis: means are equal for appliances and electronics

Customer_income Alternative Hypothesis: means are not equal for appliances and electronics

Alpha = 0.05

```{r,echo=TRUE}
ttestCI = t.test(appliances$customer_income, electronics$customer_income)
ttestCI
```

After running the t-test on appliances customer income and electronics customer income we can see that our p-value returned is 0.1725.  Comparing our p-value(0.1725) and our alpha value(0.05) we can see that the p-value is greater than alpha.  Thus, since our p-value is greater than our alpha we fail to reject the null hypothesis.  So, the evidence does not suggest that a higher or lower income is associated with purchasing electronics over appliances and vice versa.

## Outcome:

Through different hypothesis testing on these variables of store location, customer loyalty level, payment method, and customer income.  The data suggests that all of our variables are independent or share the same population mean when compared across the variable category.
