---
title: "Walmart Sales Analysis: Electronics vs Appliances (EDA)"
author: "Pablo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
```{r, echo=TRUE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
library(janitor)
library(effsize)
```

```{r, echo=TRUE}
# Make sure factors are clean
walmart <- walmart %>%
  mutate(
    category = as.factor(category),
    promotion_applied = as.factor(promotion_applied),
    promotion_type = as.factor(promotion_type)
  )
```


# Create binary category indicator for t-tests:

# 1 = Electronics, 0 = Appliances
```{r, echo=TRUE}
walmart <- walmart %>%
  mutate(category_bin = if_else(category == "Electronics", 1, 0))
```


# Graph: proportion of Electronics by promotion_applied
```{r, echo=TRUE}
walmart %>%
  group_by(promotion_applied) %>%
  summarise(prop_electronics = mean(category_bin),
            n = n(),
            .groups = "drop") %>%
  ggplot(aes(x = promotion_applied, 
             y = prop_electronics, 
             fill = promotion_applied)) +
  geom_col(width = 0.6) +
  scale_fill_manual(values = c("FALSE" = "#4E79A7",
                               "TRUE"  = "#E15759")) +
  labs(
    title = "Proportion of Electronics Transactions by Promotion Applied",
    x = "Promotion Applied",
    y = "Proportion Electronics"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

The bar chart displays the proportion of Electronics transactions among transactions with and without promotions. The proportions are nearly identical across both groups, suggesting that applying a promotion does not meaningfully shift transactions toward either Electronics or Appliances.

# t-test: does the proportion of Electronics differ by promotion_applied"

```{r, echo=TRUE}
t.test(category_bin ~ promotion_applied, data = walmart)
```

Hypotheses

Null Hypothesis (H0):
The proportion of Electronics transactions is the same for transactions with and without promotions.

Alternative Hypothesis (H1):
The proportion of Electronics transactions differs between transactions with and without promotions.


The Welch two-sample t-test shows no statistically significant difference in the proportion of Electronics transactions between transactions with and without promotions (p = 0.578). The estimated proportions are 52.1% for transactions without promotions and 51.3% for transactions with promotions. The 95% confidence interval includes zero, indicating that the true difference in proportions may be zero.

Overall, the data provide no evidence that applying a promotion changes the likelihood that a transaction belongs to the Electronics category rather than Appliances.


# Graph: quantity_sold by category
```{r, echo=TRUE}
ggplot(walmart, aes(x = category, y = quantity_sold)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_jitter(width = 0.15, alpha = 0.08) +
  labs(
    title = "Quantity Sold per Transaction by Category",
    x = "Category",
    y = "Quantity Sold"
  ) +
  theme_minimal()
```

```{r, echo=TRUE}
walmart %>%
  group_by(category) %>%
  summarise(
    mean_quantity = mean(quantity_sold),
    sd_quantity = sd(quantity_sold),
    n = n(),
    .groups = "drop"
  )
```

The average number of units purchased per transaction is nearly identical across categories. Appliances transactions have a mean of 2.97 units, while Electronics transactions have a mean of 3.00 units. The standard deviations are also very similar (1.41 and 1.43, respectively), indicating comparable variability in purchase behavior. Overall, there is no meaningful difference in the number of items purchased per transaction between the two categories.

# t-test: does mean quantity differ by category

```{r, echo=TRUE}
t.test(quantity_sold ~ category, data = walmart)
```


Null Hypothesis (H0):
The mean quantity sold per transaction is the same for Appliances and Electronics.

Alternative Hypothesis (H1):
The mean quantity sold per transaction differs between Appliances and Electronics.

The Welch two-sample t-test shows no statistically significant difference in the average number of units purchased per transaction between Appliances and Electronics (p = 0.492). Appliances transactions have a mean of 2.97 units, while Electronics transactions have a mean of 3.00 units. The 95% confidence interval includes zero, indicating that the true difference in means may be zero. Therefore, there is no statistical evidence that customers purchase more units per transaction in one category compared to the other.